---
title: "Network Analysis"
author: "VICTORIA"
date: "6/18/2024"
output:
  html_document: default
  pdf_document: default
---

## Creating and Visualizing Different Network Structures in R
Networks can be read from files and customized to specify various details, such as directionality. Here are some examples of how to create and visualize particular kinds of networks using the igraph package in R:

### Libraries needed
```{r}
library(igraph)
library(dplyr)
library(surveillance)
library(ggplot2)
library(scales)
```

### 1. Full graph
A full graph (or complete graph) is a network where every pair of distinct vertices is connected by a unique edge. This means that each node is connected to every other node.
```{r}
fg <- make_full_graph(40)
plot(fg, vertex.size=10, vertex.label=NA)
```

### 2. Star graph
A star graph consists of one central node connected to all other nodes, which are not connected to each other. It's commonly used to represent hub-and-spoke models.
```{r}
st <- make_star(40)
plot(st, vertex.size=10, vertex.label=NA)
```

### 3. Tree Graph
A tree graph is a connected acyclic graph. In this example, each node has up to 3 children. Trees are often used to represent hierarchical structures.
```{r}
tr <- make_tree(40, children = 3, mode = "undirected")
plot(tr, vertex.size=10, vertex.label=NA)
```

### 4. Ring Graph
This model generates a random graph with a fixed number of nodes (n) and edges (m). It's used to study properties of random networks.
```{r}
# Ring graph
rn <- make_ring(40)
plot(rn, vertex.size=10, vertex.label=NA)
```


### 5. Erdos-Renyi Random Graph
This model generates a random graph with a fixed number of nodes (n) and edges (m). It's used to study properties of random networks.
```{r}
# Erdos-Renyi random graph 
# ('n' is number of nodes, 'm' is the number of edges)
er <- sample_gnm(n=100, m=40) 
plot(er, vertex.size=6, vertex.label=NA)
```

### 6. Watts-Strogatz Small-World Graph
This model generates small-world networks characterized by high clustering and short path lengths. It creates a lattice structure with some edges randomly rewired.
```{r}
# Watts-Strogatz small-world graph
# Creates a lattice with 'dim' dimensions of 'size' nodes each, and rewires edges 
# randomly with probability 'p'. You can allow 'loops' and 'multiple' edges.
# The neighborhood in which edges are connected is 'nei'.
sw <- sample_smallworld(dim=2, size=10, nei=1, p=0.1)
plot(sw, vertex.size=6, vertex.label=NA, layout=layout_in_circle)
```

### 7. Barabasi-Albert preferential attachment model for scale-free graphs
This model generates scale-free networks, where the degree distribution follows a power law. New nodes preferentially attach to existing nodes with higher degrees.
```{r}
# Barabasi-Albert preferential attachment model for scale-free graphs
# 'n' is number of nodes, 'power' is the power of attachment (1 is linear)
# 'm' is the number of edges added on each time step 
ba <-  sample_pa(n=100, power=1, m=1,  directed=F)
plot(ba, vertex.size=6, vertex.label=NA)
```

### 8. Zachary Karate Club Graph
A well-known social network of a university karate club. This historical graph is often used to illustrate community detection algorithms.
```{r}
#igraph can also give you some notable historical graphs. For instance:
zach <- graph("Zachary") # the Zachary carate club
plot(zach, vertex.size=10, vertex.label=NA)

## vertex.size: Controls the size of the nodes.
## vertex.label: Determines whether to display node labels.
```

Network analysis enables us to explore the intricate connections within various types of systems. By using different graph models and visualization techniques, we can better understand the underlying structure and dynamics of these networks, providing valuable insights for research and application across multiple disciplines.